A hyperdimensional code for sequences and trees inspired by chromosomal crossover

### 2019-09-03 ms/crossing_over.tex
Include figs.

@@

\section{Introduction}
A core problem in neuroscience and artificial intelligence is how to represent a complex event like a spoken sentence using relatively fixed dimensions. Ideally similar events should have similar representations, and when a new event occurs pre-existing representations shouldn't require substantial rearrangement to prevent confusion.

@@

\State $X^Y[\mu(:, X^Y, X^{y_t}) < 1/(t+1)^\gamma] \gets X^{y_t}[\mu(:, X^Y, X^{y_t}) < 1/(t+1)^\gamma]$

@@

@@

    \For{$i \in 1, ..., M$}

        \State $\hat{X}^Y_{temp} \gets \hat{X}^Y $
        \State $\hat{X}^Y_{temp}[\mu(:, \hat{X}^Y, X^i) < 1/(t+1)^\gamma] \gets X^i[\mu(:, \hat{X}^Y, X^i) < 1/(t+1)^\gamma]$
        \State $d_{temp} \gets \textrm{Hamming}(X^Y, \hat{X}^Y_{temp})$

@@

            \State $(i_{best}, \hat{X}^Y_{best}, d_{best}) \gets (i, \hat{X}^Y_{temp}, d_{temp})$

@@

@@

@@

    \State $X^T \gets D^T[J_1]$
    \For{$k \in 2, ..., |J|$}
        \State $mask \gets k/|J| \leq \mu(:, J) \leq (k+1)/|J|$
        \State $X^T[mask] \gets J_k[mask]$
    \EndFor

@@

\State \textbf{return} $D^T[T[end]]$
\EndFunction

@@

@@

\begin{algorithmic}[0]
\Function{DecodeTree}{$Y; B, \mu$}
\State $X^Y \gets X^*$
\For{$t \in 1, ..., \textrm{len}(Y)$}

\State $X^Y[\mu(:, X^Y, X^{y_t}) < 1/(t+1)^\gamma] \gets X^{y_t}[\mu(:, X^Y, X^{y_t}) < 1/(t+1)^\gamma]$

@@

@@

This follows immediately from \ref{sec:prop-unif-comp-dstr}, since $P(x^Y_j|Y) = 1/Z$. Therefore the statistics of crossover codes do not depend on any aspect about the sequence being encoded. This distinguishes crossover, for instance, from representations that grow more or less crowded with the complexity of the sequence they represent.

>>>>

\usepackage{graphicx}

@@

\section{Introduction}
A common problem in neuroscience and artificial intelligence is how complex events, like spoken sentences, should be represented in a system of relatively fixed dimension. Ideally similar events should have similar representations, and when a new event occurs pre-existing representations shouldn't require substantial rearrangement to prevent confusion.

@@

\State $mask \gets (\mu(:, X^Y, X^{y_t}) < 1/(t+1)^\gamma)$
\State $X^Y[mask] \gets X^{y_t}[mask]$

@@

 \begin{figure}[!t]
    \centering
    \includegraphics[width=0.8\linewidth]{figs/1_encoding.png}
    \caption{Sequence encoding algorithm and properties.}
 \end{figure}

@@

    \For{$i_{temp} \in 1, ..., M$}

        \State $\hat{X}^Y_{temp} \gets \hat{X}^Y $
        \State $ mask \gets (\mu(:, \hat{X}^Y_{temp}, X^{i_{temp}}) < 1/(t+1)^\gamma)$
        \State $\hat{X}^Y_{temp}[mask] \gets X^{i_{temp}}[mask]$
        \State $d_{temp} \gets \textrm{Hamming}(X^Y, \hat{X}^Y_{temp})$

@@

            \State $(i_{best}, \hat{X}^Y_{best}, d_{best}) \gets (i_{temp}, \hat{X}^Y_{temp}, d_{temp})$

@@

\begin{figure}[!t]
    \centering
    \includegraphics[width=0.8\linewidth]{figs/2_decoding.png}
    \caption{Sequence decoding algorithm and properties.}
 \end{figure}

@@

\begin{figure}[!t]
    \centering
    \includegraphics[width=0.8\linewidth]{figs/3_gamma.png}
    \caption{Dependence of sequence codes on $\gamma$.}
 \end{figure}

@@

    \State $X^T \gets D^T[J^1]$
    \For{$k \in 2, ..., |J|$}
        \State $mask \gets (k/|J| \leq \mu(:, J) \leq (k+1)/|J|)$
        \State $X^T[mask] \gets D^T[J^k][mask]$
    \EndFor

@@

\State \textbf{return} $D^T[T[\textrm{end}]]$
\EndFunction

@@

\begin{figure}[!t]
    \centering
    \includegraphics[width=0.8\linewidth]{figs/4_trees.png}
    \caption{Tree encoding algorithm and properties.}
 \end{figure}

@@

\begin{algorithmic}[0]
\Function{DecodeTree}{$X^T; B, \mu$}
\State $D^T \gets \{i: X^i \textrm{ \textbf{if} } \textrm{Hamming}(X^T, X^i) < d_{exclude} \textrm{ \textbf{for} } i \in 1, ..., M\}$
\State $T \gets ()$

\State $d \gets 1$

    \State $(J_{best}, \hat{X}^T_{best}, d_{best}) \gets (Null, Null, 1)$

    \For{$J_{temp} \in \{J\}$}

        \State $\hat{X}^T_{temp} \gets D^T[J^1]$
        \For{$k \in 2, ..., |J_{temp}|$}
            \State $mask \gets k/|J_{temp}| \leq \mu(:, J_{temp}) \leq (k+1)/|J_{temp}|$
            \State $\hat{X}^T_{temp}[mask] \gets D^T[J^k_{temp}][mask]$
        \EndFor

        \State $d_{temp} \gets \textrm{Hamming}(X^T, \hat{X}^T_{temp})$

        \If{$d_{temp} < d_{best}$}
            \State $(J_{best}, \hat{X}^T_{best}, d_{best}) \gets (J_{temp}, \hat{X}^T_{temp}, d_{temp})$
        \EndIf
    \EndFor

    \If{$d_{best} \geq d$}
        \State \textbf{return} $T[\textrm{end}]$

    \Else
        \State $(D^T[J_{best}], d) \gets (\hat{X}^T_{best}, d_{best})$
        \State $\textrm{Append}(T, J_{best})$
    \EndIf

\State \textbf{return} $T[\textrm{end}]$
\EndFunction

@@

\begin{figure}[!t]
    \centering
    \includegraphics[width=0.8\linewidth]{figs/5_mask_corrs.png}
    \caption{Robustness to encoding noise through mask correlations.}
 \end{figure}

@@

This follows from \ref{sec:prop-unif-comp-dstr}, since $P(x^Y_j|Y) = 1/Z$. Therefore the statistics of crossover codes do not depend on any aspect about the sequence being encoded. This distinguishes crossover, for instance, from representations that grow more or less crowded with the complexity of the sequence they represent.
